Discrete Quantum Theories and Computing

### 2019-01-11 dissertation.tex
dissertation.lyx "Preface" and "Chapter 6 Further Discussion" Updating...

\usepackage{unicode-math}
\RequirePackage{ifthen}
\@ifpackageloaded{unicode-math}{}{
 \newcommand{\symbf}{\mathbf}
 \newcommand{\symrm}{\mathrm}
 \newcommand{\symcal}{\mathcal}
 \newcommand{\symbb}{\mathbb}
}
\newcommand{\muB}{\ensuremath{\mu^{\symrm{B}}}}
\newcommand{\muC}{\ensuremath{\mu^{\symrm{C}}}}
\newcommand{\muF}{\ensuremath{\mu^{\symrm{F}}}}
\newcommand{\symrmL}{\ensuremath{\symrm{L}}}
\newcommand{\symrmR}{\ensuremath{\symrm{R}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\symrmL{#1}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\symrmR{#1}}}}
\newcommand{\barmuD}{\ensuremath{\bar{\mu}^{\symrm{D}}}}
\newcommand{\events}{\ensuremath{\symcal{E}}}
\newcommand{\eventsC}{\ensuremath{\events_{\symrm{C}}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\gmult}{*}
\newcommand{\Fpx}[1]{\symbb{F}_{{#1}}}
\newcommand{\Fp}{\Fpx{p}}
\newcommand{\Fppx}[1]{\Fpx{{#1}^2}}
\newcommand{\Fpp}{\Fppx{p}}
\newcommand{\Fq}{\Fpx{q}}
\newcommand{\ff}[1]{\Fpx{#1}}
\newcommand{\ffzx}[2]{\Fpx{#1}^{{#2}\;*}}
\newcommand{\ffx}[2]{\Fpx{#1}^{{#2}}}
\newcommand{\ffzd}[1]{\ffzx{#1}{d}}
\newcommand{\ffd}[1]{\ffx{#1}{d}}
\newcommand{\VVec}[1]{\vec{\symbf{#1}}}
\newcommand{\mathReal}{\symbb{R}}
\newcommand{\mathComplex}{\symbb{C}}
\newcommand{\braket}[2]{\ip{#1}{#2}}
\newcommand{\todo}[1]{\textbf{TODO.~#1}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Sphere}[1]{\symbf{S}^{#1}}
\newcommand{\R}{\mathReal}
\newcommand{\rme}{\symrm{e}}
\newcommand{\rmi}{\symrm{i}}
\newcommand{\rmd}{\symrm{d}}
\newcommand{\CP}[1]{\mathComplex\symbf{P}^{#1}} % Complex projective space
\newcommand{\DCP}[1]{\symbf{D}\mathComplex\symbf{P}^{#1}} % Discrete complex projective
\def\fh{\mathfrak{h}}
\newcommand{\uf}{U_{\!f}}
\newcommand{\scalarPlus}{+}
\newcommand{\boolt}{\textsf{Bool}} 
\newcommand{\bfalse}{\texttt{\textbf{false}}}
\newcommand{\btrue}{\texttt{\textbf{true}}}
\newcommand{\dotprod}{dot product}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\newcommand{\Psibar}{\overline{\Psi}}
\newcommand{\cardp}[2]{#1 \, \slash\!\! \slash \, #2}
\def\round{\mathop{\rm round}\nolimits}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\Hilb}{\symcal{H}}
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\def\C{{\symbb{C}}}
\newcommand{\ultramodular}{\symcal{M}}
\newcommand{\ultramodularL}[1][]{\ensuremath{\ultramodular^{L{#1}}}}
\newcommand{\ultramodularR}[1][]{\ensuremath{\ultramodular^{R{#1}}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\nb}{\nolinebreak[3] }

@@

\begin{preface*}
\begin{itemize}
\item The latest version of this thesis is in GitHub: \code{https://git.io/fxbuG}
\item Its source code is typeset using \LyX ~\cite{LyX} in: \code{https://git.io/fxbuE}
\item Any comments can be left in the Issue part of the GitHub Repository:
\code{https://git.io/fxbug}
\end{itemize}
\end{preface*}

@@

\subsection{Ultramodular Functions}

@@

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states, and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
As we discussed in the previous section, the finite-precision extension
of the Gleason theorem, Thm.~\ref{thm:Finite-precision-Gleason},
only applies on the QIVPMs composted by interval maps $f:\left[0,1\right]\rightarrow\mathscr{I}$
and quantum probability measures $\muB_{\rho}:\events\rightarrow\left[0,1\right]$,
where the maximum length of intervals in $\mathscr{I}$ is an important
. In this case, the states consistent with the composition $f\circ\muB_{\rho}$
is bounded by . While the QIVPM defined in Table~\ref{tab:non-Born-QIVPM}
is not in this class, it is too wild to have a non-empty core, and
the maximum length of its intervals $\mathscr{I}_{0}$ is $1$, the
longest length of all possible intervals. This suggests us to ask
if we shirk the maximum length of intervals alone, whether this ill-behaved
QIVPMs disappears. More precisely, we have the following question.

\begin{question}In a Hilbert space $\Hilb$ of dimension $D\geq3$,
given a QIVPM $\bar{\mu}:\events\rightarrow\mathscr{I}$, $\alpha$
be the maximum length of intervals in $\mathscr{I}$, is there an
$\varepsilon>0$ such that $\coreBorn\left(\bar{\mu},\events\right)\ne\emptyset$
for $\alpha<\varepsilon$.\end{question}

\noindent %
\begin{comment}

@@

\todo{Add a two-page conclusion chapter which could include further
discussion, but it can be only a few sentences? }\todo{Discrete
Schrödinger equation? And its relation with Grover's algorithm? }\todo{Notice
that in differential geometry, Killing–Hopf theorem asserts that having
positive constant curvature guarantees a manifold is essentially a
sphere, but say little things even if the curvature has only small
deviation from constant. Since a quantum probability measure is glued
by classical probability measures, like a manifold. When their relations
are exact, Gleason's theorem asserts that a quantum probability measure
can essentially induced by the Born rule, but they can be wild when
we move on to interval-valued situation! }
\end{comment}

>>>>

\usepackage{unicode-math}
\RequirePackage{ifthen}
\@ifpackageloaded{unicode-math}{}{
 \newcommand{\symbf}{\mathbf}
 \newcommand{\symrm}{\mathrm}
 \newcommand{\symcal}{\mathcal}
 \newcommand{\symbb}{\mathbb}
}
\newcommand{\muB}{\ensuremath{\mu^{\symrm{B}}}}
\newcommand{\muC}{\ensuremath{\mu^{\symrm{C}}}}
\newcommand{\muF}{\ensuremath{\mu^{\symrm{F}}}}
\newcommand{\symrmL}{\ensuremath{\symrm{L}}}
\newcommand{\symrmR}{\ensuremath{\symrm{R}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\symrmL{#1}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\symrmR{#1}}}}
\newcommand{\barmuD}{\ensuremath{\bar{\mu}^{\symrm{D}}}}
\newcommand{\events}{\ensuremath{\symcal{E}}}
\newcommand{\eventsC}{\ensuremath{\events_{\symrm{C}}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\gmult}{*}
\newcommand{\Fpx}[1]{\symbb{F}_{{#1}}}
\newcommand{\Fp}{\Fpx{p}}
\newcommand{\Fppx}[1]{\Fpx{{#1}^2}}
\newcommand{\Fpp}{\Fppx{p}}
\newcommand{\Fq}{\Fpx{q}}
\newcommand{\ff}[1]{\Fpx{#1}}
\newcommand{\ffzx}[2]{\Fpx{#1}^{{#2}\;*}}
\newcommand{\ffx}[2]{\Fpx{#1}^{{#2}}}
\newcommand{\ffzd}[1]{\ffzx{#1}{d}}
\newcommand{\ffd}[1]{\ffx{#1}{d}}
\newcommand{\VVec}[1]{\vec{\symbf{#1}}}
\newcommand{\mathReal}{\symbb{R}}
\newcommand{\mathComplex}{\symbb{C}}
\newcommand{\braket}[2]{\ip{#1}{#2}}
\newcommand{\todo}[1]{\textbf{TODO.~#1}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Sphere}[1]{\symbf{S}^{#1}}
\newcommand{\R}{\mathReal}
\newcommand{\rme}{\symrm{e}}
\newcommand{\rmi}{\symrm{i}}
\newcommand{\rmd}{\symrm{d}}
\newcommand{\CP}[1]{\mathComplex\symbf{P}^{#1}} % Complex projective space
\newcommand{\DCP}[1]{\symbf{D}\mathComplex\symbf{P}^{#1}} % Discrete complex projective
\def\fh{\mathfrak{h}}
\newcommand{\uf}{U_{\!f}}
\newcommand{\scalarPlus}{+}
\newcommand{\boolt}{\textsf{Bool}} 
\newcommand{\bfalse}{\texttt{\textbf{false}}}
\newcommand{\btrue}{\texttt{\textbf{true}}}
\newcommand{\dotprod}{dot product}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\newcommand{\Psibar}{\overline{\Psi}}
\newcommand{\cardp}[2]{#1 \, \slash\!\! \slash \, #2}
\def\round{\mathop{\rm round}\nolimits}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\Hilb}{\symcal{H}}
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\def\C{{\symbb{C}}}
\newcommand{\ultramodular}{\symcal{M}}
\newcommand{\ultramodularL}[1][]{\ensuremath{\ultramodular^{L{#1}}}}
\newcommand{\ultramodularR}[1][]{\ensuremath{\ultramodular^{R{#1}}}}
\newcommand{\frameL}[1][]{\ensuremath{f^{L{#1}}}}
\newcommand{\frameR}[1][]{\ensuremath{f^{R{#1}}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\nb}{\nolinebreak[3] }

@@

\begin{preface*}
\begin{itemize}
\item The latest PDF version of this thesis is in GitHub: \code{https://git.io/fxbuG}.
\item Its source code is typeset using \LyX ~\cite{LyX} in \code{https://git.io/fxbuE}
and the exported \LaTeX{} file is in \code{https://git.io/fhZBW}.
\item Any comments can be left in the Issue part of the GitHub repository:
\code{https://git.io/fxbug}.
\end{itemize}
Following explains how to compile the PDF from the \LaTeX{} source:
\begin{enumerate}
\item \label{enu:Clone-the-git}Clone the git repository of this thesis
\code{https://git.io/fhZP5} into your local directory and the thesis
document class \code{https://github.com/yuttai/iuphd} in a location
where the system can find.
\item In the local directory of this thesis, run \code{xelatex dissertation.tex}.
\item Run \code{biber dissertation}.
\item Run \code{xelatex dissertation.tex} several times until all the references
are resolved.
\end{enumerate}
The previous steps only tested in Microsoft Windows. Please report
me if there is any difficulty compile it in the other system. Besides,
to compile the PDF file from \LyX{} source, just finish step~\ref{enu:Clone-the-git}
and click the toolbar button \includegraphics[width=1em]{buffer-view}.\end{preface*}

@@

\subsection{Ultramodular Functions\label{subsec:Ultramodular-Functions}}

@@

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states, and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
As we discussed in the previous section, the finite-precision extension
of the Gleason theorem, Thm.~\ref{thm:Finite-precision-Gleason},
only applies on the QIVPMs composted by interval maps $f:\left[0,1\right]\rightarrow\mathscr{I}$
and quantum probability measures $\muB_{\rho}:\events\rightarrow\left[0,1\right]$;
the states consistent with the composition $f\circ\muB_{\rho}$ converges
to $\rho$ as the maximum length of intervals in $\mathscr{I}$ shrinking
to $0$. In contrast, the wildest QIVPM defined in Table~\ref{tab:non-Born-QIVPM}
maps to $\mathscr{I}_{0}$ whose intervals' maximum length is $1$,
the longest possible length. This provides another possibility that
shirking the maximum length of intervals itself could result QIVPMs
with non-empty cores. More precisely, we have the following question.

\begin{question}In a Hilbert space $\Hilb$ of dimension $D\geq3$,
given a QIVPM $\bar{\mu}:\events\rightarrow\mathscr{I}$ with $\alpha$
being the maximum length of intervals in $\mathscr{I}$, is there
an $\varepsilon>0$ such that $\coreBorn\left(\bar{\mu},\events\right)\ne\emptyset$
for $\alpha<\varepsilon$.\end{question}

@@

\todo{Add a two-page conclusion chapter which could include further
discussion, but it can be only a few sentences? }\todo{Discrete
Schrödinger equation? And its relation with Grover's algorithm? }\todo{Notice
that in differential geometry, Killing–Hopf theorem asserts that having
positive constant curvature guarantees a manifold is essentially a
sphere, but say little things even if the curvature has only small
deviation from constant. Since a quantum probability measure is glued
by classical probability measures, like a manifold. When their relations
are exact, Gleason's theorem asserts that a quantum probability measure
can essentially induced by the Born rule, but they can be wild when
we move on to interval-valued situation! }
\end{comment}
To better understand this question, we start from the dimension $D=3$.
This case is simpler than the general situation because any two-dimensional
projectors can be expressed as the complement of a one-dimensional
projector, $\mathbb{1}-\ket{\psi}$. Together with Eq.~\ref{eq:QIVPM-complement},
the interval-valued probability of any two-dimensional projectors,
$\bar{\mu}\left(\mathbb{1}-\ket{\psi}\right)$, can also be expressed
as the complement of the latter, $\necess-\bar{\mu}\left(\ket{\psi}\right)$.
Since the one-dimensional projectors are one-to-one corresponding
to the irreducible states in the complex projective space $\CP{2}$
as discussed in Sec.~\ref{subsec:Explicit-generalization-of}, QIVPMs
$\bar{\mu}:\events\rightarrow\mathscr{I}$ are one-to-one corresponding
to interval-valued frame functions $\bar{f}:\CP{2}\rightarrow\mathscr{I}$
according to the following lemma.

\begin{lemma}\label{thm:interval-valued-frame-function}Given a QIVPM
$\bar{\mu}:\events\rightarrow\mathscr{I}$, the function $\bar{f}:\CP{2}\rightarrow\mathscr{I}$
by $\bar{f}\left(\ket{\psi}\right)=\bar{\mu}\left(\proj{\psi}\right)$
are called an interval-valued frame function, and satisfies
\begin{equation}
\left[1,1\right]-\bar{f}\left(\ket{\psi_{0}}\right)\subseteq\bar{f}\left(\ket{\psi_{1}}\right)+\bar{f}\left(\ket{\psi_{2}}\right)\label{eq:interval-valued-frame-function}
\end{equation}
for any orthonormal basis $\left\{ \ket{\psi_{i}}\right\} _{i=0}^{2}$.
Conversely, if a function $\bar{f}:\CP{2}\rightarrow\mathscr{I}$
satisfies Eq.~\ref{eq:interval-valued-frame-function}, $\bar{f}$
must be an interval-valued frame function.\end{lemma}

\begin{proof}This lemma is the direct consequence of Lemma~\ref{thm:convex-3-1}.\end{proof}

Similar to the ultramodular functions in Sec.~\ref{subsec:Ultramodular-Functions},
we can also decompose an interval-valued frame function into its left-end
and right-end, where $\left[\frameL\left(\ket{\psi}\right),\frameR\left(\ket{\psi}\right)\right]=\bar{f}\left(\ket{\psi}\right)$.
Then, Eq.~\ref{eq:interval-valued-frame-function} can be rephrased
as
\begin{equation}
\left[1,1\right]-\bar{f}\left(\ket{\psi_{0}}\right)\subseteq\bar{f}\left(\ket{\psi_{1}}\right)+\bar{f}\left(\ket{\psi_{2}}\right)\label{eq:two-ends-frame-function}
\end{equation}
